---
title: "1. XGBoost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# XGBoost

## Fit model with xgboost library

```{r}
source("read/read_xgboost_norandom.R")
library(xgboost)
library(caret)
```

http://datasideoflife.com/?p=1009

## Prepare data

```{r}
x_train <- as.matrix(train %>% 
                       dplyr::select(weeks, months, years))
x_test <- as.matrix(test %>% 
                      dplyr::select(weeks, months, years))
y_train <- train$incid
```

## Grid search

```{r}
xgb_trcontrol <- caret::trainControl(
  method = "cv", 
  number = 10,
  allowParallel = TRUE, 
  verboseIter = FALSE, 
  returnData = FALSE,
  search = "grid"
)
```

```{r}
set.seed(1234)
xgb_grid <- expand.grid(max_depth = c(4, 6, 10, 15), 
                        nrounds = c(1500, 2000, 5000, 7000),    # number of trees
                        # default values below
                        eta = 0.2,
                        gamma = 0,
                        subsample = 0.01,
                        min_child_weight = 0.01,
                        colsample_bytree =0.01)
```

## Model

```{r}
# xgb_model <- caret::train(
#   incid ~ weeks+months+years,
#   data=train,
#   trControl = xgb_trcontrol,
#   tuneGrid = xgb_grid,
#   method = "xgbTree",
#   nthread = 5
# )
```

```{r}
# xgb_model$bestTune
```

```{r}
# xgb_test <- xgb_model %>% stats::predict(x_test, se.fit=TRUE)
```

## Plots

```{r}
# df_plot <- as.data.frame(cbind(extended_data[109:161, ], xgb_test))
# df_plot2 <- df_plot %>% rename(Original = incid,
#                                  Predicted = ...7)
# 
# library(ggplot2)
# # png(file="figures/xgboost_norandom.png",
# #     width=850, height=500)
# ggplot(data = df_plot2, aes(x=date)) +
#   geom_line(aes(y=Original, color="Original"))+
#   geom_line(aes(y=Predicted, color="Predicted"))+
#   labs(x="Weeks", y="Weekly cases per 100 000")+
#   scale_color_manual(name="Data", values = c(
#     "Original"="black", 
#     "Predicted"="red"))+ 
#   theme_minimal()
# # dev.off()
```

## Error measurements

```{r}
# measures=function(pred,obs){
#   
#   err=pred-obs
#   perr=err/pred
#   
#   c(RMSE=sqrt(mean(err^2)),
#     MAE=mean(abs(err)),
#     nMSE=sqrt(mean(err^2))/mean(pred),
#     nMAE=mean(abs(err))/mean(pred),
#     RMSPE=sqrt(mean(perr^2)),
#     MAPE=mean(abs(perr)))
# }
```

```{r}
# error_test <- measures(xgb_test, test$incid)
# error_test
```

########


## Predictions with test data

Run the model 500 times and obtain 500 predictions for the 53 weeks in the test dataset. 

```{r}
# Results
load(file="read/1. XGBoost-predictions500.RData")
```

Los siguientes 2 chunks no hace falta ejecutarlos. Los resultados se cargan en el chunk anterior. 

```{r, eval=FALSE}
predictions <- data.frame(matrix(0,53,500))
set.seed(123)
seeds <- runif(500,1,100000)
```

```{r, eval=FALSE}
for(i in 1:ncol(predictions)){
  set.seed(seeds[i])
  xgb_model <- caret::train(
  incid ~ weeks+months+years,
  data=train,
  trControl = xgb_trcontrol,
  tuneGrid = xgb_grid,
  method = "xgbTree",
  nthread = 5)

  predictions[, i] <- xgb_model %>% stats::predict(x_test, se.fit=TRUE)
}
```

### Mean points

```{r}
predictions_median <- apply(as.matrix(predictions), 1, function(x){quantile(x, c(0.5))})
```

```{r}
df_plot <- as.data.frame(cbind(extended_data[109:161, ],predictions_median))
df_plot2 <- df_plot %>% rename(Original = incid,
                                 Median_Predicted = ...7)

ggplot(data = df_plot2, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black", 
    "Median_Predicted"="red"))+ 
  theme_minimal()
```

### Credible interval   

```{r}
CI <- apply(as.matrix(predictions), 1, function(x){quantile(x, c(0.025, 0.5, 0.975))})
```

```{r}
CI_t <- as.data.frame(t(CI))
colnames(CI_t) <- c("CI_low", "Median", "CI_upper")
```

```{r}
xgboost500 <- cbind(df_plot2, CI_t)
```

```{r}
library(ggplot2)
# png(file="figures/xgboost_500.png",
#     width=850, height=500)
ggplot(data = xgboost500, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  geom_ribbon(aes(ymin=CI_low, ymax=CI_upper), alpha = 0.1)+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black",
    "Median_Predicted"="red"))+
  theme_minimal()
# dev.off()
```

### Error measurements

```{r}
measures=function(pred,obs){

  err=pred-obs
  perr=err/pred

  c(RMSE=sqrt(mean(err^2)),
    MAE=mean(abs(err)),
    nMSE=sqrt(mean(err^2))/mean(pred),
    nMAE=mean(abs(err))/mean(pred),
    RMSPE=sqrt(mean(perr^2)),
    MAPE=mean(abs(perr)))
}
```

```{r}
error_test <- measures(xgboost500$Median_Predicted, xgboost500$Original)
error_test
```


########


## Predictions with train data

Run the model 500 times and obtain 500 predictions for the 108 weeks in the train dataset. 

```{r}
# Results
load(file="read/1. XGBoost-predictions500_train.RData")
```

Los siguientes 2 chunks no hace falta ejecutarlos. Los resultados se cargan en el chunk anterior. 

```{r, eval=FALSE}
predictions <- data.frame(matrix(0,53,500))
set.seed(123)
seeds <- runif(500,1,100000)
```

```{r, eval=FALSE}
for(i in 1:ncol(predictions)){
  set.seed(seeds[i])
  xgb_model <- caret::train(
  incid ~ weeks+months+years,
  data=train,
  trControl = xgb_trcontrol,
  tuneGrid = xgb_grid,
  method = "xgbTree",
  nthread = 5)

  predictions[, i] <- xgb_model %>% stats::predict(x_train, se.fit=TRUE)
}
```

### Mean points

```{r}
predictions_median <- apply(as.matrix(predictions), 1, function(x){quantile(x, c(0.5))})
```

```{r}
df_plot <- as.data.frame(cbind(extended_data[1:108, ],predictions_median))
df_plot2 <- df_plot %>% rename(Original = incid,
                                 Median_Predicted = ...7)

ggplot(data = df_plot2, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black", 
    "Median_Predicted"="blue"))+ 
  scale_x_date(date_labels =  "%b %Y") +
  theme_minimal()
```

### Credible interval   

```{r}
CI <- apply(as.matrix(predictions), 1, function(x){quantile(x, c(0.025, 0.5, 0.975))})
```

```{r}
CI_t <- as.data.frame(t(CI))
colnames(CI_t) <- c("CI_low", "Median", "CI_upper")
```

```{r}
xgboost500_train <- cbind(df_plot2, CI_t)
```

```{r}
library(ggplot2)
# png(file="figures/xgboost_500_train.png",
#     width=850, height=500)
ggplot(data = xgboost500_train, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  geom_ribbon(aes(ymin=CI_low, ymax=CI_upper), alpha = 0.1)+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black",
    "Median_Predicted"="blue"))+
  scale_x_date(date_labels =  "%b %Y") +
  theme_minimal()
# dev.off()
```

### Error measurements

```{r}
measures=function(pred,obs){

  err=pred-obs
  perr=err/pred

  c(RMSE=sqrt(mean(err^2)),
    MAE=mean(abs(err)),
    nMSE=sqrt(mean(err^2))/mean(pred),
    nMAE=mean(abs(err))/mean(pred),
    RMSPE=sqrt(mean(perr^2)),
    MAPE=mean(abs(perr)))
}
```

```{r}
error_train <- measures(xgboost500_train$Median_Predicted, xgboost500_train$Original)
error_train
```