---
title: "2. Neuralnet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Neural Net

https://datascienceplus.com/neuralnet-train-and-test-neural-networks-using-r/

## Data 

```{r}
library(neuralnet)
source("read/read_neuralnet.R")
```

## Prepare time series

```{r}
m <- 53 # coger todo el año -> explicar las dinámicas anuales
ny <- 1
mi <- min(cases_tot$incid)*0.9
ma <- max(cases_tot$incid)*1.1

dades <- ts(cases_tot$incid)
for (i in 1:m){
  dades <- ts.intersect(dades,stats::lag(cases_tot$incid,-i))
}

dades <- as.data.frame(dades)
names(dades) <- paste0("y",c("",paste0("_",1:m)))
dades <- (dades-mi)/(ma-mi)
```

## Neural network

```{r}
kf <- 1
npred <- 53
idx <- c(rep(1,nrow(dades)-npred),rep(2,npred)) #definir indice

train <- as.matrix(dades[idx!=2,])
test <- as.matrix(dades[idx==2,])

resp <- "y"
if(ny>1) resp <- paste0(resp,c("",paste0("_",1:(ny-1))),collapse="+")
expl <- paste0("y_",ny:m, collapse = "+")
f <- as.formula(paste0(resp, "~",expl))
```

```{r}
set.seed(123456)
net <- neuralnet(f, train, lifesign="full", lifesign.step =5000, hidden = c(7,5), rep=1, threshold=0.00125)

plot(net) #saved as 950*800
```

## Predictions

```{r}
pred.train <- mi+predict(net,train[nrow(train)-(npred-1):0,-c(1:ny)])*(ma-mi)
pred.test <- mi+predict(net,test[,-c(1:ny)])*(ma-mi)
```

```{r}
obs.train <- mi+train[nrow(train)-(npred-1):0, 1:ny]*(ma-mi) #train data
obs.test <- mi+test[,1:ny]*(ma-mi) #test data
```

## Plot

```{r}
df_plot <- as.data.frame(cbind(cases_tot[109:161, ], pred.test))
df_plot2 <- df_plot %>% rename(Original = incid,
                              Predicted = pred.test)

library(ggplot2)
# png(file="figures/neuralnet_test.png",
#     width=850, height=500)
ggplot(data = df_plot2, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Predicted, color="Predicted"))+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black", 
    "Predicted"="red"))+ 
  theme_minimal()
# dev.off()
```

## Error measurements

```{r}
measures=function(pred,obs){
  
  err=pred-obs
  perr=err/pred
  
  c(RMSE=sqrt(mean(err^2)),
    MAE=mean(abs(err)),
    nMSE=sqrt(mean(err^2))/mean(pred),
    nMAE=mean(abs(err))/mean(pred),
    RMSPE=sqrt(mean(perr^2)),
    MAPE=mean(abs(perr)))
}
```

```{r}
error_train <- measures(pred.train, obs.train)
error_train
```

```{r}
error_test <- measures(pred.test, obs.test)
error_test
```

########

## Predictions with test data

Run the model 500 times and obtain 500 predictions for the 53 weeks in the test dataset. 

```{r}
# Results
load(file="read/2. Neuralnet-predictions500.RData")
```

Los siguientes 2 chunks no hace falta ejecutarlos. Los resultados se cargan en el chunk anterior. 

```{r}
predictions <- vector("list", 500)
set.seed(123)
seeds <- runif(500,1,100000)
```

```{r}
for(i in 1:length(predictions)){
  set.seed(seeds[i])
  net <- neuralnet(f, train, lifesign="full", lifesign.step =5000, hidden = c(7,5), rep=1, threshold=0.00125)
  
  pred.test <-mi+predict(net,test[,-c(1:ny)])*(ma-mi) #test
  predictions[[i]] <- pred.test
}
```

### Median points

```{r}
n <- length(predictions[[1]])
predictions_df <- structure(predictions, row.names = c(NA, -n), .Names = seq_along(predictions), class = "data.frame")
```

```{r}
predictions_median <- apply(as.matrix(predictions_df), 1, function(x){quantile(x, c(0.5))})
```

```{r}
df_plot <- as.data.frame(cbind(cases_week[109:161, ],predictions_median))
df_plot2 <- df_plot %>% rename(Original = incid,
                                 Median_Predicted = ...5)

ggplot(data = df_plot2, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black", 
    "Median_Predicted"="red"))+ 
  theme_minimal()
```


### Credible interval   

```{r}
CI <- apply(as.matrix(predictions_df), 1, function(x){quantile(x, c(0.025, 0.5, 0.975))})
```

```{r}
CI_t <- as.data.frame(t(CI))
colnames(CI_t) <- c("CI_low", "Median", "CI_upper")
```

```{r}
neuralnet500 <- cbind(df_plot2, CI_t)
```

```{r}
library(ggplot2)
# png(file="figures/neuralnet_500.png",
#     width=850, height=500)
ggplot(data = neuralnet500, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  geom_ribbon(aes(ymin=CI_low, ymax=CI_upper), alpha = 0.1)+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black",
    "Median_Predicted"="red"))+
  theme_minimal()
# dev.off()
```

### Error measurements

```{r}
measures=function(pred,obs){

  err=pred-obs
  perr=err/pred

  c(RMSE=sqrt(mean(err^2)),
    MAE=mean(abs(err)),
    nMSE=sqrt(mean(err^2))/mean(pred),
    nMAE=mean(abs(err))/mean(pred),
    RMSPE=sqrt(mean(perr^2)),
    MAPE=mean(abs(perr)))
}
```

```{r}
error_test <- measures(neuralnet500$Median_Predicted, neuralnet500$Original)
error_test
```
########


## Predictions with train data

Run the model 500 times and obtain 500 predictions for 53 weeks in the train dataset. 

```{r}
# Results
load(file="read/2. Neuralnet-predictions500_train.RData")
```

Los siguientes 2 chunks no hace falta ejecutarlos. Los resultados se cargan en el chunk anterior. 

```{r}
predictions <- vector("list", 500)
set.seed(123)
seeds <- runif(500,1,100000)
```

```{r}
for(i in 1:length(predictions)){
  set.seed(seeds[i])
  net <- neuralnet(f, train, lifesign="full", lifesign.step =5000, hidden = c(7,5), rep=1, threshold=0.00125)
  
  pred.train <- mi+predict(net,train[nrow(train)-(npred-1):0,-c(1:ny)])*(ma-mi) #train
  predictions[[i]] <- pred.train
}
```

### Median points

```{r}
n <- length(predictions[[1]])
predictions_df <- structure(predictions, row.names = c(NA, -n), .Names = seq_along(predictions), class = "data.frame")
```

```{r}
predictions_median <- apply(as.matrix(predictions_df), 1, function(x){quantile(x, c(0.5))})
```

```{r}
df_plot <- as.data.frame(cbind(cases_week[55:107, ],predictions_median)) #first 53
df_plot2 <- df_plot %>% rename(Original = incid,
                                 Median_Predicted = ...5)

ggplot(data = df_plot2, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black", 
    "Median_Predicted"="blue"))+ 
  theme_minimal()
```


### Credible interval   

```{r}
CI <- apply(as.matrix(predictions_df), 1, function(x){quantile(x, c(0.025, 0.5, 0.975))})
```

```{r}
CI_t <- as.data.frame(t(CI))
colnames(CI_t) <- c("CI_low", "Median", "CI_upper")
```

```{r}
neuralnet500_train <- cbind(df_plot2, CI_t)
```

```{r}
library(ggplot2)
# png(file="figures/neuralnet_500_train.png",
#     width=850, height=500)
ggplot(data = neuralnet500_train, aes(x=date)) +
  geom_line(aes(y=Original, color="Original"))+
  geom_line(aes(y=Median_Predicted, color="Median_Predicted"))+
  geom_ribbon(aes(ymin=CI_low, ymax=CI_upper), alpha = 0.1)+
  labs(x="Weeks", y="Weekly cases per 100 000")+
  scale_color_manual(name="Data", values = c(
    "Original"="black",
    "Median_Predicted"="blue"))+
  theme_minimal()
# dev.off()
```

### Error measurements

```{r}
measures=function(pred,obs){

  err=pred-obs
  perr=err/pred

  c(RMSE=sqrt(mean(err^2)),
    MAE=mean(abs(err)),
    nMSE=sqrt(mean(err^2))/mean(pred),
    nMAE=mean(abs(err))/mean(pred),
    RMSPE=sqrt(mean(perr^2)),
    MAPE=mean(abs(perr)))
}
```

```{r}
error_train <- measures(neuralnet500_train$Median_Predicted, neuralnet500_train$Original)
error_train
```